\bf Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test

### 2025-06-10 manuscript/manuscript.tex
cleaned up figures

We now consider the case where $X_i$'s are realizations from a time series and
$X^{(b)}_1, \cdots ,X^{(b)}_n$ are generated by block bootstrap for
$b \in \{1, \ldots, B\}$.
Circular block-bootstrap can be done with overlapping blocks that can
wrap around the end of the series \citet{romano1992circular}.
Define blocks (assuming block size $l > 1$) as:
\begin{equation*}
Z_j =
    \begin{cases}
        \{X_j, \ldots, X_{j + l - 1}\}, & j = 1, \dots, n - l + 1,\\
        \{X_j, \ldots, X_n, X_1, \ldots, X_{j-n+l-1}\}, & j = n - l
        + 2 ,\dots, n.
    \end{cases}
\end{equation*}
A common
function for block size that is considered optimal is
$l = \lceil n^{1/3} \rceil$ \citep{buhlmann1999block}.
\citet{hall1995blocking} and
\citet{politis2004automatic} note that
the selection of block size should ideally take into account the strength
and direction of serial dependence. The plug-in approach for circular
bootstrap introduced in \citet{politis2004automatic}.
Now we draw $k$ blocks from the $(n - l + 1)$ blocks
of $Z_j$'s with replacement and then align them in the order they were picked to
form a block bootstrap sample. If $n$ is not a multiple of~$l$, the last block
selected will be reduced in size so that the final size of the block bootstrap
sample is $n$.

@@

These distributions
were chosen to compare results on normal and non-normal
error structures. The specific parameters were chosen because the distributions
are very similar and their
first two moments are the same.
For the purposes of
evaluating if the test holds it size, when $X_i \sim N(8, 8)$, we tested that the
marginal distribution is from the
Normal family, or
$X_i \sim N(\cdot ; \mu, \sigma^2), \quad i = 1, \ldots, n$
for some unspecified $\mu$ and $\sigma$. When $X_i \sim \Gamma(8, 1)$, we tested
that the marginal distribution family is from the Gamma family,
or
$X_i \sim \Gamma(\cdot ; \alpha, \beta), \quad i = 1, \ldots, n$.
for some unspecified $\alpha$ and $\beta$. For the block bootstrap step,
we used $B = 1000$ and $l = \lceil n^{1/3} \rceil$.
For each setting for $F$, $\tau$, and $n$, we replicate the method 10000 times
to get the distribution of the p-values $p_r$ for
$r \in \{1, \dots, 10000\}$
for the test when applied to samples from the same data generating process.

@@

In our original setting, we used a block size of $l = \lceil n^{1/3} \rceil$.
However, the optimal block size may vary according to the dependence structure.
To address this, we considered  using the automatic block-length selection
scheme
for
circular block bootstrap as described in Section 3.3 of
\citet{politis2004automatic}. Unfortunately, using this setting results in
the selection of block
sizes that can be quite large (sometimes as large or larger than $n$), leading
to less variation between the bootstrapped
samples. Figures~\ref{fig:large_block}
demonstrate that for highly temporally
dependent samples (especially when the autocorrelation is negative),
the block size is frequently larger than $\sqrt{n}$.

@@

Even for settings where the block sizes are usually less than $\sqrt{n}$ -
such as for $n = 800$, $\tau = 0.5$, and $X_i \sim N(8, 8)$ - the 
block
sizes are almost always larger than $l = \lceil n^{1/3} \rceil$.

\jy{How do you interpret the flat lines in the plot? Why are the
  empirical cdf step functions? That means your empirical p-values are
  discrete with point masses at a few points. Check the p-values from
  the simulations.}
\mc{this was a result of erroneous data, it is fixed now}

\begin{figure}[tbp]
  \centering
  \includegraphics[scale=1]{figures/block_dist}
  \caption{Distribution of the selected block sizes using
  \citet{politis2004automatic}'s procedure applied to 20000 length-800 time 
  series with
    Kendall's $\tau = 0.5$ data generating distribution $N(8,8)$.
  }
  \label{fig:block_dist}
\end{figure}

Despite this, our 
algorithm holds its size somewhat well for moderately dependent
series as demonstrated in 
Figures~\ref{fig:alt_normal} and \ref{fig:alt_gamma}. Both 
\citet{politis2004automatic}'s setting and the cubic root setting result
in the method not holding its size for $|\tau| > 0.5$. 
Figure~\ref{fig:alt_rr} also shows that the two settings are comparably
powerful for $|\tau| <= 0.5$. However, we still
recommend the cubic root setting over that introduced by 
\citet{politis2004automatic} because the latter fails to hold its size
for negative $\tau$, particularly for smaller sample sizes. It is important to 
note that another block size selection rationale could yield better results.

\begin{figure}[tbp]
  \centering
  \includegraphics[scale=1]{figures/alt_normal}
  \caption{Q-Q plots of the p-values testing that a time series
    have marginal normal distribution with true data generating distribution
    being $N(8,8)$, when using \citet{politis2004automatic}'s procedure
    to select block sizes.
  }
  \label{fig:alt_normal}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[scale=1]{figures/alt_gamma}
  \caption{Q-Q plots of the p-values testing that a time series
    have marginal gamma distribution with true data generating distribution
    being $\Gamma(8,1)$, when using \citet{politis2004automatic}'s procedure
    to select block sizes.
  }
  \label{fig:alt_gamma}
\end{figure}

>>>>

We now consider the case where $X_i$'s are realizations from a time series and
$X^{(b)}_1, \cdots ,X^{(b)}_n$ are generated by block bootstrap for
$b \in \{1, \ldots, B\}$.
Circular block-bootstrap can be done with overlapping blocks that can
wrap around the end of the series \citep{romano1992circular}.
Define blocks (assuming block size $l > 1$) as:
\begin{equation*}
Z_j =
    \begin{cases}
        \{X_j, \ldots, X_{j + l - 1}\}, & j = 1, \dots, n - l + 1,\\
        \{X_j, \ldots, X_n, X_1, \ldots, X_{j-n+l-1}\}, & j = n - l
        + 2 ,\dots, n.
    \end{cases}
\end{equation*}
A common
function for block size that is often considered optimal is
$l = \lceil n^{1/3} \rceil$ \citep{buhlmann1999block},
although
\citet{hall1995blocking} and
\citet{politis2004automatic} note that
the selection of block size should ideally take into account the strength
and direction of serial dependence. 
Now we draw $k$ blocks from the $(n - l + 1)$ blocks
of $Z_j$'s with replacement and then align them in the order they were picked to
form a block bootstrap sample. If $n$ is not a multiple of~$l$, the last block
selected will be reduced in size so that the final size of the block bootstrap
sample is $n$.

@@

These distributions
were chosen to compare results on normal and non-normal
error structures. The specific parameters were chosen because the distributions
are very similar and their
first two moments are the same.
For the purposes of
evaluating if the test holds it size, when $X_i \sim N(8, 8)$, we tested that the
marginal distribution is from the
Normal family, or
$X_i \sim N(\cdot ; \mu, \sigma^2), \quad i = 1, \ldots, n$
for some unspecified $\mu$ and $\sigma$. When $X_i \sim \Gamma(8, 1)$, we tested
that the marginal distribution family is from the Gamma family,
or
$X_i \sim \Gamma(\cdot ; \alpha, \beta), \quad i = 1, \ldots, n$.
for some unspecified $\alpha$ and $\beta$. For the block bootstrap step,
we used $B = 1000$ and $l = \lceil n^{1/3} \rceil$.
As noted in Subsection~\ref{sub:npbb}, previous literature
suggests varying block size based on the dependence structure of the series.
For this reason,
we also tried implementing our algorithm using the plug-in approach for circular
bootstrap block size introduced in \citet{politis2004automatic}. Results and
a discussion of this additional simulation are provided in the Appendix.
For each setting for $F$, $\tau$, and $n$, we replicate the method 10000 times
to get the distribution of the p-values $p_r$ for
$r \in \{1, \dots, 10000\}$
for the test when applied to samples from the same data generating process.

@@

In our original setting, we used a block size of $l = \lceil n^{1/3} \rceil$.
However, the optimal block size may vary according to the dependence structure.
To address this, we considered  using the automatic block-length selection
scheme
for
circular block bootstrap as described in Section 3.3 of
\citet{politis2004automatic}. Unfortunately, using this setting results in
the selection of block
sizes that can be quite large (sometimes as large or larger than $n$), leading
to less variation between the bootstrapped
samples. Figures~\ref{fig:large_block}
demonstrate that for highly temporally
dependent samples (especially when the autocorrelation is negative),
the block size is frequently larger than $\sqrt{n}$. For more moderate 
positive temporal dependence, the block size may not be overly large
as frequently.

@@

Still, even for settings where the block sizes are usually less than $\sqrt{n}$ -
such as for $n = 800$, $\tau = 0.5$, and $X_i \sim N(8, 8)$ - the 
block sizes are almost always larger than $l = \lceil n^{1/3} \rceil$, a setting
which was considered optimal by \citep{buhlmann1999block}. This
is demonstrated in Figure~\ref{fig:block_dist} with a histogram
of the selected block sizes for this setting.

\begin{figure}[tbp]
  \centering
  \includegraphics[scale=1]{figures/block_dist}
  \caption{Distribution of the selected block sizes using
  \citet{politis2004automatic}'s procedure applied to 20000 length-800 time 
  series with
    Kendall's $\tau = 0.5$ data generating distribution $N(8,8)$. The vertical
    dashed line is $l = \lceil n^{1/3} \rceil$ for $n = 800$, which is equivalent 
    to 10.
  }
  \label{fig:block_dist}
\end{figure}

Despite this, the
algorithm with \citet{politis2004automatic}'s setting holds its size somewhat 
well for moderately 
positively dependent
series as demonstrated in 
Figure~\ref{fig:alt_qq}. Both 
\citet{politis2004automatic}'s setting and the cubic root setting result
in the method not holding its size for $|\tau| > 0.5$. 
Figure~\ref{fig:alt_rr} also shows that the two settings are comparably
powerful for $|\tau| <= 0.5$, with \citet{politis2004automatic}'s
setting perhaps even having higher rejection rates for smaller negatively
dependent sample sizes. 
However, we still
recommend the original cubic root setting over that introduced by 
\citet{politis2004automatic} because the latter fails to hold its size
for negative $\tau$, particularly for smaller sample sizes.

\begin{figure}[tbp]
  \includegraphics[width = .9\textwidth]{figures/alt_normal}
  \centering
  \vspace{-10pt}
  \caption{Q-Q plots of the p-values testing that a time series
    have marginal normal distribution with true data generating distribution
    being $N(8,8)$, when using \citet{politis2004automatic}'s procedure
    to select block sizes.}
    \hspace{3cm}
  \includegraphics[width = .9\textwidth]{figures/alt_gamma}
  \vspace{-5pt}
  \caption{Q-Q plots of the p-values testing that a time series
    have marginal gamma distribution with true data generating distribution
    being $\Gamma(8,1)$, when using \citet{politis2004automatic}'s procedure
    to select block sizes.}
  \label{fig:alt_qq}
\end{figure}

### 2025-06-10 manuscript/reply.tex
typo fix
cleaned up figures

\jy{Make sure you've included all the comments from the reviewers,
  including their openning summaries.}

@@

\begin{comment}
Two reviewers have carefully read the paper and have provided detailed feedback.  Both reviewers would like to see some additional discussion and justifications for the bias correction term, where one could closely follow the Babu and Rao (2004) paper, and one reviewer mentioned the importance of considering different block sizes (I recommend possibly trying some data-based block size selection method and checking how it performs).  A successful revision will require that the authors give serious consideration to these points.
\end{comment}

@@

\jy{Is there a second point from the AE?}

@@

\jy{Make sure all references here are cited at appropriate places in the paper.}

@@

\begin{comment}
This paper focuses on approximating the distribution of the Kolmogorov-
Smirnov (KS) statistic for stationary time series. Specifically, it deals with the
distribution of KS statistic for testing if the marginal distribution belongs to
a specified parametric family. Parametric bootstrap requires knowledge of the
nature of serial dependence, and for this reason, nonparametric bootstrap is 
employed here. For the iid case, \citet{babu2004goodness} obtained a correction 
term so that the corrected bootstraped statistic can be used to approximate 
distribution of the KS statistic. This paper does the same except it uses 
nonparametric bootstrap developed by Kunsch and others. Finally the author(s) 
present some empirical work in support of their proposed procedure. Simulations 
indicate that the proposed procedure may work for large samples and moderate depen-
dence.
\end{comment}

@@

\begin{comment}
1. In Section 3.1, it seems that Kendall's tau is being used as a measure
dependence between $X_t$ and $X_{t+1}$. It may be useful to make that explicit 
in the second pararaph of the Size subsection.
\end{comment}

@@

Here is what I am able to understand after reading this paper, the papers
by Babu and Rao, and Kunsch and others.

Following the notations in this paper (the same notations have been used in
Babu and Rao), let

\mc{I will insert the rest of the comment here}
\end{comment}

>>>>

@@

\begin{comment}
Two reviewers have carefully read the paper and have provided detailed feedback.  Both reviewers would like to see some additional discussion and justifications for the bias correction term, where one could closely follow the \citet{babu2004goodness} paper, and one reviewer mentioned the importance of considering different block sizes (I recommend possibly trying some data-based block size selection method and checking how it performs).  A successful revision will require that the authors give serious consideration to these points.
\end{comment}

@@

@@

@@

\begin{comment}
This paper focuses on approximating the distribution of the Kolmogorov-
Smirnov (KS) statistic for stationary time series. Specifically, it deals with the
distribution of KS statistic for testing if the marginal distribution belongs to
a specified parametric family. Parametric bootstrap requires knowledge of the
nature of serial dependence, and for this reason, nonparametric bootstrap is 
employed here. For the iid case, \citet{babu2004goodness} obtained a correction 
term so that the corrected bootstraped statistic can be used to approximate 
distribution of the KS statistic. This paper does the same except it uses 
nonparametric bootstrap developed by Kunsch and others. Finally the author(s) 
present some empirical work in support of their proposed procedure. Simulations 
indicate that the proposed procedure may work for large samples and
moderate dependence.
\end{comment}

@@

\begin{comment}
1. In Section 3.1, it seems that Kendall's tau is being used as a measure
dependence between $X_t$ and $X_{t+1}$. It may be useful to make that explicit 
in the second paragraph of Section 3.1.
\end{comment}

@@

Here is what I am able to understand after reading this paper, the papers
by \citet{babu2004goodness}, and \citet{kunsch1989jackknife} and others.

Following the notations in this paper (the same notations have been used in
\citet{babu2004goodness}), let

\begin{align*}
Y_n(x, \hat \theta_n) &= \sqrt{n}[F_n(x) - F(x; \hat \theta_n)],\\
Y_n^{(b)}(x, \hat \theta_n^{(b)}) &= \sqrt{n}[F_n^{(b)}(x) - F(x; \hat \theta_n^{(b)})],
\end{align*}

where the second quantity is based on the $b$th bootstrap sample. In the iid 
case, \citet{babu2004goodness} introduced a correction term for the nonparametric bootstrap
$$B_n(x) = \sqrt{n}[F_n(x) - F(x; \hat \theta_n)],$$
so that
\begin{align*}
Y_{n,corr}^{(b)}(x) &= Y_n^{(b)}(x; \hat\theta_n^{(b)}) - B_n(x)\\
                    &= \sqrt{n}[F_n^{(b)}(x) - F_n(x)] -
                    \sqrt{n}[F(x; \hat \theta_n^{(b)}) -  F(x; \hat\theta_n)].
\end{align*}
It was proved in \citet{babu2004goodness} that the distribution of the KS
statistic is approximated by that of the corrected bootstrap under the null as
well as for contiguous alternatives.

The work under consideration uses nonparametric block bootstrap. Let 
$\theta^*_n = E^*[\hat \theta_n ^{(b)}]$ and $F^*_n(x) = E^*[F_n^{(b)}(x)]$,
where $E^*$ represents average over bootstrap samples. In this paper the 
correction term is
$$C_n(x) = \sqrt{n}[F^*_n(x)-F(x;\theta^*_n)],$$
so that
\begin{align*}
Y^{(b)}_{n,corr}(x) &= Y^{(b)}_n(x;\hat \theta_n^{(b)})-C_n(x)\\
                    &= \sqrt{n}[F_n^{(b)}(x) - F^*_n(x)] 
                    - \sqrt{n}[F(x; \hat\theta_n^{(b)}) - F(x; \theta^*_n)].
\end{align*}
So the corrected bootstrap proposed here has the same form as in Babu and Rao,
and the justifications for the procedure proposed here should also be similar.
\end{comment}
